### 2024-07-11 doc/manuscript.Rmd
methods

---
title: "Soil drivers of plant, fungal and bacterial community novelty in urban habitats of the city of Xixón, Spain"
output: 
  bookdown::github_document2:
    reference_docx: template.docx
    number_sections: false
bibliography: refs.bib
csl: journal-of-vegetation-science.csl
link-citations: yes
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",")
})
```

Running title: Urban drivers of plant, fungal and bacterial novelty

Eduardo Fernández-Pascual^1,a^ (<https://orcid.org/0000-0002-4743-9577>), Zuzana Ferencova^2^ (<https://orcid.org/0000-0002-0915-0916>),
Víctor González-García^1^ (<https://orcid.org/0000-0002-8949-7943>), Borja Jiménez-Alfaro^1^ (<https://orcid.org/0000-0001-6601-9597>)

^1^ Biodiversity Research Institute (IMIB), University of Oviedo - CSIC - Principality of Asturias, Mieres, Spain; ^2^ Jardín Botánico Atlántico, Gijón/Xixón, Spain

Correspondence: Eduardo Fernández-Pascual, Instituto Mixto de Investigación en Biodiversidad, Campus de Mieres, Edificio de Investigación, 5ª planta, c/ Gonzalo Gutiérrez Quirós s/n, E-33600 Mieres, Spain. Email: [fernandezpeduardo\@uniovi.es](mailto:fernandezpeduardo@uniovi.es){.email}. Telephone: +34985104781.

# Author contributions

Eduardo Fernández-Pascual: Conceptualization; Methodology; Investigation; Data Curation; Formal Analysis; Visualization; Writing – Original Draft Preparation; Writing – Review & Editing. Zuzana Ferencova: Conceptualization; Methodology; Investigation; Data Curation; Writing – Review & Editing. Víctor González-García: Investigation; Borja Jiménez-Alfaro: Methodology; Writing – Review & Editing.

# Acknowledgements

María García Fernández, Sara Gutiérrez Cruz and Abelardo Sigüenza Solís helped with the field surveys.

# Funding

Fundación Biodiversidad, Ministerio para la Transición Ecológica y el Reto Demográfico, Gobierno de España [XIXÓN ECORRESILIENTE, PRTR, NextGenerationEU]. Ayuntamiento de Gijón/Xixón, Jardín Botánico Atlántico [SV-23-GIJON-JBA].

# Conflict of interest statement

None declared.

# Data availability statement

Original datasets, R code for analysis and Rmarkdown code for creation of the manuscript can be accessed at Zenodo: https://zenodo.org/doi/10.5281/zenodo.11613571.

# Introduction

In post-industrial landscapes affected by the phenomenon of urban contraction, land located in the heart of cities remains unused for long periods (Haase et al., 2014) . When the use of any urban plot ceases, it is colonized by plants from the regional flora capable of living in urban environments (Johnson et al., 2017) . These urban plants are adapted to disturbances and can grow in degraded soils; some even accumulate heavy metals, making them nature-based solutions for bioremediation of degraded lands (Song et al., 2019) . Species traits, along with the legacy of human impact, determine which regional flora species colonize which plots (Swan et al., 2021) . When the human legacy is intense, the plant communities that develop will be novel communities, different from the historical reference communities that existed before human intervention (Gedan and Fernández-Pascual, 2019) . The novelty of the communities can be reflected in their floristic, compositional and phylogenetic composition, including their proportion of exotic species (Lososová et al., 2012) . In any case, proper management of urban communities can turn them into resources with high ecological and social potential (Anderson and Minor, 2017). (Kowarik, 2018) : in the Ibero-Atlantic ecoregion, anthropogenic habitats can host up to a third of the ecoregional flora.

The ability of plants to colonize any habitat, including urban ones, depends largely on their seed regeneration niche, that is, on their dispersal, persistence, germination and establishment functions (Grubb, 1977, Saatkamp et al., 2019) . However, although it is known that urban species often share a series of functional and ecological traits (Aronson et al., 2007, Lososová et al., 2006) , little is known about their seed regeneration niche. Especially, information on its germination traits is lacking. In species living in non-anthropogenic environments, germination phenology is regulated by environmental factors (water, temperature, light) to ensure that germination occurs at the right time to maximize subsequent seedling survival (Carta et al., 2022a) . Compared to the large amount of data on the germination of species from non-anthropogenic habitats, little is known about what factors regulate the germination of urban species. Considering other habitats affected by humans (Fernández-Pascual et al., 2022) , it could be expected that urban seeds have their germination unrelated to macroclimatic signals and show tolerance to stress due to heat, drought or heavy metals. However, this hypothesis needs to be tested through a synthesis of the seed regeneration niche in urban environments.

Cities 
1. highly fragmented landscapes
2. mosaics of heterogeneous ecosystems (in terms of size, fragmentation, population density and land use)
3. subject to different degrees of human interference, leading to a stepwise transformation of pristine ecosystems to novel urban ecosystems
4. comprised of a patchwork of paved and unpaved spaces, built and vacant land, and newly developed and obsolescent and/or abandoned buildings and infrastructure
5. contains a multitude of vacant lots, railway sidings, utility easements, corridors between buildings and canal sides that are often overgrown with spontaneous vegetation, which are not coherently managed, and which seem to occupy an uncertain, interstitial niche in the urban matrix

Cities are comprised of a patchwork of paved and unpaved spaces, built and vacant land, and newly developed and obsolescent and/or abandoned buildings and infrastructure. They are mosaics of heterogeneous ecosystems (in terms of size, fragmentation, population density and land use) which are subject to different degrees of human interference, leading to a stepwise transformation of pristine ecosystems to novel urban ecosystems (Kowarik, 2005).

The “Four Nature approach” narrows down the variety of transformational stages in urban settings to four major types that can supply wilderness in urban region (Kowarik 1992, 2005). Each of the four types relates to nature in general but results from different trajectories in human-nature interactions:

Spontaneous vegetation appears in different habitat types in the urban matrix,. Multitude of vacant lots, railway sidings, utility easements, corridors between buildings and canal sides that are often overgrown with spontaneous vegetation, which are not coherently managed, and which seem to occupy an uncertain, interstitial niche in the urban matrix. The spontaneous herbs could provide the very interesting pool of species for ecological restoration of the urban area. Their ecological value has been clear since many years because their contribution to the ecosystem services in urban areas. Although their aesthetic value has been highlighted repeatedly, their acceptance by the residents in urban areas has been questioned. Many residents associate the presence of spontaneous plants with the neglect in maintenance by public (municipal) management service. Spontaneous native urban vegetation has commonly been described as demonstrating resilience (Ignatieva et al., 2000) and exhibiting adaptations to human disturbance (Lundholm and Marlin, 2006; Sukopp, 2004).

Urban species pool is comprised of native species present before the urban development, native species developed because the new urban ecosystem and introduced species by human activity. However, just some of them can adapt to urban environment.

IGS (informal green space) consists of any urban space with a history of strong anthropogenic disturbance that is covered at least partly with non-remnant, spontaneous vegetation (Del Tredici, 2010).

# Materials and methods

## Study city

Gijón/Xixón is a medium-sized coastal city (population c. 270,000 inhabitants) located in the Bay of Biscay, within the central Asturias metropolitan area of northern Spain (population c. 800,000 inhabitants). The city belongs to the Cantabrian Mixed Forests ecoregion [@RN4667], at the border between the temperate and Mediterranean climatic zones of Europe, and has a temperate oceanic climate with cool summers and wet and mostly frost-free winters (mean annual temperature = 13.5 ºC, annual precipitation = 1050 mm). Most of the city is built on alluvial and estuarine deposits, sand, or limestone bedrock. The potential pre-human vegetation would be dominated by deciduous mixed forests, together with riparian forests, wetlands and coastal dunes. The area has a long history of human habitation dating back to the Cantabrian Upper Paleolithic [@RN5450]. The city itself has been inhabited at least since Roman times [@RN5532]. During the medieval and early modern era the urban center was limited to the original walled city, with fishing as the main economic activity. Expansion beyond the original core started in the 18th and 19th centuries. Exposure to global trade networks and the mild climate turned the region into a hotspot for biological invasions, especially from the Americas and Asia [@RN5493]. In 1852, the city was connected by railway and became a major port for the regional mining and industrial centers. During the 20th century the city experienced significant industrial development (e.g. canning, coal, steel, shipbuilding) and rapid population growth, becoming the most populated city in the Principality of Asturias [@RN5534; @RN5533]. Industrial and urban growth brought environmental problems including loss of natural wetlands and air and watercourse pollution [@RN5535]. Like the rest of its metropolitan area, the city has been affected by recent post-industrial land-use changes which have left large areas with abandoned industrial sites (i.e. brownfields) [@RN5448; @RN5449].

## Definition of urban habitats

To survey urban biodiversity across habitats, we defined six habitats types (**Figure 1A**) representing different states within the four-nature classification of urban wilderness [@RN5445]: (1) forest patches < 10 km from the city center, i.e. type-1 nature in remnants of the pre-urban vegetation; (2) hay meadows, i.e. type-2 nature in remnants of human-made ecosystems resulting from agricultural habitat transformation; (3) parks, i.e. type-3 nature in managed green spaces created after habitat destruction; (4) roadsides, i.e. type-4 nature in relatively unmanaged strips of land alongside roads, railways, etc; (5) residential lots, i.e. type-4 nature in unmanaged vacant lots of previous residential use; and (6) industrial lots, i.e. type-4 nature in unmanaged vacant lots of previous  industrial use.

## Selection of sampling plots

During June 2023, we surveyed the city to establish five 4 x 4 m sampling plots per each of the six habitat types (n = 30 plots). The plots were distributed across the city (**Figure 1B**) and represented the variation of vegetation types found within each habitat type. In two plots per habitat (**Figure 1B**), we buried a datalogger (M-Log5W, GeoPrecision, Ettlingen, Germany; accuracy: +/- 0.1 ºC (at 0 ºC), resolution: 0.01 ºC) at a depth of 5 cm below the upper layer of the soil. Dataloggers recorded temperature once every hour and stayed on site for approximately one year, after which we retrieved them and downloaded their records to describe the thermal environment of the city and the habitats.

## Soil physicochemical sampling

From each of the 30 plots, we collected 500 g of soil for analysis of chemical composition and physical soil properties. We took the samples from 5 points in each plot (center + 4 cardinal points). We removed stones and other debris particles, sieved the soil in a 5 mm mesh and homogenized the sample in a plastic bag. We cleaned and sterilized all tools with water and ethanol between samples. We sent the samples for analysis at INEA - Escuela Universitaria de Ingeniería Agrícola, Universidad Pontificia de Comillas. Soil analysis included pH, texture, bulk density, electrical conductivity, organic matter content, total nitrogen, C/N ratio, active limestone, total carbonates, assimilable phosphorus, K, Ca, Mg, Na and heavy metals (Zn, Cr, Cu, Cd, Hg, Pb, Ni).

> Añadir detalles de los métodos de suelo (Zuzana)

## Plant, fungal and bacterial diversity sampling

In each plot, we recorded all the spontaneous vascular plant species, as well as the cover (%) of each species in the plot. For plant species names we followed Euro+Med [@RN5044]. Additionally, from each of the 30 plots, we collected soil samples for DNA metabarcoding analysis of soil fungal and bacterial communities. We performed the sampling in sterile conditions, i.e. before entering the plot for any of the other tasks. In each plot, we took samples from the four cardinal points, 0.5 m inwards from the edge, at a depth of 5 cm. Before taking the samples, we removed the vegetation with a sterile tool. Then, we collected the soil samples in a 50 ml Falcon tube, mixed them, homogenized them with a lancet and finally stored the sample into a 2 ml Eppendorf tube which we immediately froze. We sent the frozen tube to AllGenetics & Biology SL (www.allgenetics.eu) for DNA metabarcoding analysis.

## DNA metabarcoding analysis

DNA analysis by AllGenetics included DNA extraction; library preparation with 2 pairs of different primers specific for barcoding (ITS for fungi and 16S for bacteria); sequencing on an Illumina NovaSeq PE250 platform (assuming an amplicon size < 450 bp); and standard bioinformatic analysis including demultiplexing, quality control and sample preprocessing, taxonomic assignment, number of reads per taxon and generation of rarefaction curves.

*DNA isolation and quantification*. DNA was isolated from each soil sample using the DNeasy PowerSoil Pro DNA isolation kit (Qiagen) following the manufacturer’s instructions, and DNA was eluted in a final volume of 50 μL. In one sample the above-mentioned protocol did not yield a DNA extract of enough concentration, so DNA was isolated using the FastDNA Spin Kit for Soil (MPBio) following the manufacturer’s instructions, and DNA was eluted in a final volume of 70 μL. An extraction blank (Bex) was included in each round of DNA extraction, and treated as the plot samples to check for contamination. DNA concentration in each extract was quantified using the Qubit High Sensitivity dsDNA Assay (Thermo Fisher Scientific).

*DNA metabarcoding library preparation and sequencing*. For fungal library preparation, a fragment of the nuclear ITS1 region of around 280-300 bp (including the primer sequences) was amplified using the following primers: forward - ITS1f (5’ CTTGGTCATTTAGAGGAAGTAA 3’) [@RN5548]; reverse - ITS2 (5’ GCTGCGTTCTTCATCGATGC 3’) [@RN5549]. For prokaryotic library preparation, a fragment of the 16S rRNA gene of around 300 bp (including primer sequences) was amplified using the following primers: forward - 515F-Y (5’ GTGYCAGCMGCCGCGGTAA 3’) [@RN5536]; reverse - 806RB (5’ GGACTACNVGGGTWTCTAAT 3’) [@RN5537]. These primers also included the Illumina sequencing primer sequences attached to their 5’ ends. In the first amplification step, PCRs were carried out in a final volume of 12.5 μL, containing 1.25 μL of template DNA (1:2 diluted in PCR round 3; not diluted in PCR round 11), 0.5 μM of the primers, 3.13 μL of Supreme NZYTaq 2x Green Master Mix (NZYTech), and ultrapure water up to 12.5 μL. The fungal reaction mixture was incubated as follows: an initial denaturation at 95 ºC for 5 min, followed by 35 cycles of 95 ºC for 30 s, 49 ºC for 45 s, 72 ºC for 45 s, and a final extension step at 72 ºC for 7 min. The bacterial reaction mixture was incubated as follows: an initial denaturation at 95 ºC for 5 min, followed by 25 cycles of 95 ºC for 30 s, 47 ºC for 45 s, 72ºC for 45 s, and a final extension step at 72 ºC for 7 min. The oligonucleotide indices that are required for multiplexing different libraries in the same sequencing pool were attached in a second amplification step with identical conditions but only 5 cycles and 60 ºC as the annealing temperature. A negative control that contained no DNA (BPCR) was included in every PCR round to check for contamination during library preparation. Library size was verified by running the libraries on 2 % agarose gels stained with GreenSafe (NZYTech) and imaging them under UV light. Then, libraries were purified using the Mag-Bind RXNPure Plus magnetic beads (Omega Bio-tek), following the instructions provided by the manufacturer. Finished libraries were pooled in equimolar amounts according to the results of a Qubit dsDNA HS Assay (Thermo Fisher Scientific) quantification. The pool was sequenced in a fraction of a NovaSeq PE250 flow cell (Illumina) aiming for a total output of 2 gigabases. Any potential traces of adapter dimers were removed using Cutadapt v3.5 [@RN5538]. Quality of the sequencing data was assesed with the software FastQC [@RN5539].

*Processing of sequencing data and inference of ASVs*. Cutadapt v3.5 was used to trim off non-biological DNA sequences (primers, indices, and sequencing adapters). Then, amplicon reads were processed using the DADA2 pipeline [@RN5541] implemented in QIIME 2 (release 2023.7) [@RN5540]. The first step in the DADA2 pipeline was to remove PCR primers and filter reads according to their quality. After checking the read quality profiles, forward reads were truncated at position 175, and reverse reads at position 131. The next step was to denoise and infer Amplicon Sequence Variants (ASVs). Error rates were learned from the dataset to denoise, using the parametric error model implemented in DADA2. Dereplication of the dataset (i.e. combination of all identical reads into unique reads) was carried out to reduce computational effort. These dereplicated forward and reverse reads were used to infer ASVs with the core sample inference algorithm [@RN5541]. Subsequently, corresponding R1 and R2 reads were merged into pairs with a minimum overlap of 12 identical base pairs. The final step in the DADA2 pipeline was to reduce the impact of artefacts (chimaeras) in the dataset. The final output of the DADA2 pipeline was a table containing the count of reads of every observed ASV in each sample, as well as a table listing all ASVs and their corresponding representative sequences.

*Taxonomic assignment*. Taxonomic assignment of each ASV was performed using a pre-trained classifier of the UNITE reference database for fungi [@RN5550] (release May 2021); and of the SILVA reference database for bacteria [@RN5542] (release 138.1 August 2020). To compare the representative sequences of the ASVs to the reference database, and compute the taxonomic assignment,the algorithm sklearn [@RN5543] implemented in QIIME 2 was applied, setting a minimum confidence level of 0.8. Several filters were applied to exclude: (1) singletons (i.e. ASVs containing only one member sequence in the whole data set); (2) ASVs occurring at a frequency below 0.01 % in each sample, to reduce the mistagging bias [@RN5544]; (3) from the bacterial dataset, non-prokaryotic ASVs such as eukaryotic sequences of plastid [@RN5545] and mitochondrial [@RN5546] origin; (4) those assigned only at domain (Bacteria) or kingdom (Fungi) level; (5) ASVs present in the negative controls that were more abundant in the negative controls than in the samples, indicating contamination [@RN5547].

*Alpha rarefaction plots*. Alpha rarefaction plots showed that the sequencing depth to reach saturation was 30,000 reads for the fungi and 22,000 reads for the bacteria, above the sequencing depth of all plots for fungi (mean = 114,919 reads; min = 79,045; max = 157,110) and bacteria (mean = 113,341 reads; min = 84,094; max = 171,404).

> Rarefaction

## Statistical analysis

# Results

## Soil temperatures and properties across habitats

## Plant, fungal and bacterial diversity across habitats

> PERMANOVAs

> Analysis with species, genera and ASVs

> Analysis with presence absence

## Community novelty across habitats and taxonomic groups

## Soil drivers of novelty

# Discussion

The recorded flora represents just a subset of possible spontaneous vegetation because this study was carried on during limited time which covers just part of the growing season of the whole vegetation.

There are some other habitats in the city which were not included in the present study, e.g. wetlands, beach, cliffs.