### 2024-07-15 doc/manuscript.Rmd
first draft of methods and results

---
title: "Soil drivers of plant, fungal and bacterial community novelty in urban habitats of the city of Xixón, Spain"
output: 
  bookdown::github_document2:
    reference_docx: template.docx
    number_sections: false
bibliography: refs.bib
csl: journal-of-vegetation-science.csl
link-citations: yes
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",")
})
```

Running title: Urban drivers of plant, fungal and bacterial novelty

Eduardo Fernández-Pascual^1,a^ (<https://orcid.org/0000-0002-4743-9577>), Zuzana Ferencova^2^ (<https://orcid.org/0000-0002-0915-0916>),
Víctor González-García^1^ (<https://orcid.org/0000-0002-8949-7943>), Borja Jiménez-Alfaro^1^ (<https://orcid.org/0000-0001-6601-9597>)

^1^ Biodiversity Research Institute (IMIB), University of Oviedo - CSIC - Principality of Asturias, Mieres, Spain; ^2^ Jardín Botánico Atlántico, Gijón/Xixón, Spain

Correspondence: Eduardo Fernández-Pascual, Instituto Mixto de Investigación en Biodiversidad, Campus de Mieres, Edificio de Investigación, 5ª planta, c/ Gonzalo Gutiérrez Quirós s/n, E-33600 Mieres, Spain. Email: [fernandezpeduardo\@uniovi.es](mailto:fernandezpeduardo@uniovi.es){.email}. Telephone: +34985104781.

# Author contributions

Eduardo Fernández-Pascual: Conceptualization; Methodology; Investigation; Data Curation; Formal Analysis; Visualization; Writing – Original Draft Preparation; Writing – Review & Editing. Zuzana Ferencova: Conceptualization; Methodology; Investigation; Data Curation; Writing – Review & Editing. Víctor González-García: Investigation; Borja Jiménez-Alfaro: Methodology; Writing – Review & Editing.

# Acknowledgements

María García Fernández, Sara Gutiérrez Cruz and Abelardo Sigüenza Solís helped with the field surveys.

# Funding

Fundación Biodiversidad, Ministerio para la Transición Ecológica y el Reto Demográfico, Gobierno de España [XIXÓN ECORRESILIENTE, PRTR, NextGenerationEU]. Ayuntamiento de Gijón/Xixón, Jardín Botánico Atlántico [SV-23-GIJON-JBA].

# Conflict of interest statement

None declared.

# Data availability statement

Original datasets, R code for analysis and Rmarkdown code for creation of the manuscript can be accessed at Zenodo: https://zenodo.org/doi/10.5281/zenodo.11613571.

# Introduction

In post-industrial landscapes affected by the phenomenon of urban contraction, land located in the heart of cities remains unused for long periods (Haase et al., 2014) . When the use of any urban plot ceases, it is colonized by plants, fungi or bacteria from the regional species pools that capable of living in urban environments (Johnson et al., 2017) . This urban biota is adapted to disturbance and can grow in degraded soils; this should make these species useful for nature-based solutions such as bioremediation of degraded lands (Song et al., 2019) . Species traits, along with the legacy of human impact, determine which regional pool species colonize which plots (Swan et al., 2021) . When the human legacy is intense, the communities that develop will be novel communities, different from the historical reference communities that existed before human intervention (Gedan and Fernández-Pascual, 2019) . The novelty of the communities can be reflected in their composition, including their elements such as the proportion of exotic species or the proportion of obligate synanthropic species (Lososová et al., 2012) . In any case, proper management of urban communities can turn them into resources with high ecological and social potential (Anderson and Minor, 2017). (Kowarik, 2018). Cities are comprised of a patchwork of paved and unpaved spaces, built and vacant land, and newly developed and obsolescent and/or abandoned buildings and infrastructure. They are mosaics of heterogeneous ecosystems (in terms of size, fragmentation, population density and land use) which are subject to different degrees of human interference, leading to a stepwise transformation of pristine ecosystems to novel urban ecosystems (Kowarik, 2005). The “Four Nature approach” narrows down the variety of transformational stages in urban settings to four major types that can supply wilderness in urban region (Kowarik 1992, 2005).

# Materials and methods

## Study city

Gijón/Xixón is a medium-sized coastal city (population c. 270,000 inhabitants) located in the Bay of Biscay, within the central Asturias metropolitan area of northern Spain (population c. 800,000 inhabitants). The city belongs to the Cantabrian Mixed Forests ecoregion [@RN4667], at the border between the temperate and Mediterranean climatic zones of Europe, and has a temperate oceanic climate with cool summers and wet and mostly frost-free winters (mean annual temperature = 13.5 ºC, annual precipitation = 1050 mm). Most of the city is built on alluvial and estuarine deposits, sand, or limestone bedrock. The potential pre-human vegetation would be dominated by deciduous mixed forests, together with riparian forests, wetlands and coastal dunes. The area has a long history of human habitation dating back to the Cantabrian Upper Paleolithic [@RN5450]. The city itself has been inhabited at least since Roman times [@RN5532]. During the medieval and early modern era the urban center was limited to the original walled city, with fishing as the main economic activity. Expansion beyond the original core started in the 18th and 19th centuries. Exposure to global trade networks and the mild climate turned the region into a hotspot for biological invasions, especially from the Americas and Asia [@RN5493]. In 1852, the city was connected by railway and became a major port for the regional mining and industrial centers. During the 20th century the city experienced significant industrial development (e.g. canning, coal, steel, shipbuilding) and rapid population growth, becoming the most populated city in the Principality of Asturias [@RN5534; @RN5533]. Industrial and urban growth brought environmental problems including loss of natural wetlands and air and watercourse pollution [@RN5535]. Like the rest of its metropolitan area, the city has been affected by recent post-industrial land-use changes which have left large areas with abandoned industrial sites (i.e. brownfields) [@RN5448; @RN5449].

## Definition of urban habitats

To survey urban biodiversity across habitats, we defined six habitats types (**Figure 1A**) representing different states within the four-nature classification of urban wilderness [@RN5445]: (1) forest patches < 10 km from the city center, i.e. type-1 nature in remnants of the pre-urban vegetation; (2) hay meadows, i.e. type-2 nature in remnants of human-made ecosystems resulting from agricultural habitat transformation; (3) parks, i.e. type-3 nature in managed green spaces created after habitat destruction; (4) roadsides, i.e. type-4 nature in relatively unmanaged strips of land alongside roads, railways, etc; (5) residential lots, i.e. type-4 nature in unmanaged vacant lots of previous residential use; and (6) industrial lots, i.e. type-4 nature in unmanaged vacant lots of previous  industrial use.

## Selection of sampling plots

During June 2023, we surveyed the city to establish five 4 x 4 m sampling plots per each of the six habitat types (n = 30 plots). The plots were distributed across the city (**Figure 1B**) and represented the variation of vegetation types found within each habitat type. In two plots per habitat (**Figure 1B**), we buried a datalogger (M-Log5W, GeoPrecision, Ettlingen, Germany; accuracy: +/- 0.1 ºC (at 0 ºC), resolution: 0.01 ºC) at a depth of 5 cm below the upper layer of the soil. Dataloggers recorded temperature once every hour and stayed on site for approximately one year, after which we retrieved them and downloaded their records to describe the thermal environment of the city and the habitats.

## Soil physicochemical sampling

From each of the 30 plots, we collected 500 g of soil for analysis of chemical composition and physical soil properties. We took the samples from 5 points in each plot (center + 4 cardinal points). We removed stones and other debris particles, sieved the soil in a 5 mm mesh and homogenized the sample in a plastic bag. We cleaned and sterilized all tools with water and ethanol between samples. We sent the samples for analysis at INEA - Escuela Universitaria de Ingeniería Agrícola, Universidad Pontificia de Comillas. Soil analysis included pH, texture, bulk density, electrical conductivity, organic matter content, total nitrogen, C/N ratio, active limestone, total carbonates, assimilable phosphorus, K, Ca, Mg, Na and heavy metals (Zn, Cr, Cu, Cd, Hg, Pb, Ni).

> Añadir detalles de los métodos de suelo (Zuzana)

## Plant, fungal and bacterial diversity sampling

In each plot, we recorded all the spontaneous vascular plant species, as well as the cover (%) of each species in the plot. For plant species names we followed Euro+Med [@RN5044]. Additionally, from each of the 30 plots, we collected soil samples for DNA metabarcoding analysis of soil fungal and bacterial communities. We performed the sampling in sterile conditions, i.e. before entering the plot for any of the other tasks. In each plot, we took samples from the four cardinal points, 0.5 m inwards from the edge, at a depth of 5 cm. Before taking the samples, we removed the vegetation with a sterile tool. Then, we collected the soil samples in a 50 ml Falcon tube, mixed them, homogenized them with a lancet and finally stored the sample into a 2 ml Eppendorf tube which we immediately froze. We sent the frozen tube to AllGenetics & Biology SL (www.allgenetics.eu) for DNA metabarcoding analysis.

## DNA metabarcoding analysis

DNA analysis by AllGenetics included DNA extraction; library preparation with 2 pairs of different primers specific for barcoding (ITS for fungi and 16S for bacteria); sequencing on an Illumina NovaSeq PE250 platform (assuming an amplicon size < 450 bp); and standard bioinformatic analysis including demultiplexing, quality control and sample preprocessing, taxonomic assignment, number of reads per taxon and generation of rarefaction curves.

*DNA isolation and quantification*. DNA was isolated from each soil sample using the DNeasy PowerSoil Pro DNA isolation kit (Qiagen) following the manufacturer’s instructions, and DNA was eluted in a final volume of 50 μL. In one sample the above-mentioned protocol did not yield a DNA extract of enough concentration, so DNA was isolated using the FastDNA Spin Kit for Soil (MPBio) following the manufacturer’s instructions, and DNA was eluted in a final volume of 70 μL. An extraction blank (Bex) was included in each round of DNA extraction, and treated as the plot samples to check for contamination. DNA concentration in each extract was quantified using the Qubit High Sensitivity dsDNA Assay (Thermo Fisher Scientific).

*DNA metabarcoding library preparation and sequencing*. For fungal library preparation, a fragment of the nuclear ITS1 region of around 280-300 bp (including the primer sequences) was amplified using the following primers: forward - ITS1f (5’ CTTGGTCATTTAGAGGAAGTAA 3’) [@RN5548]; reverse - ITS2 (5’ GCTGCGTTCTTCATCGATGC 3’) [@RN5549]. For prokaryotic library preparation, a fragment of the 16S rRNA gene of around 300 bp (including primer sequences) was amplified using the following primers: forward - 515F-Y (5’ GTGYCAGCMGCCGCGGTAA 3’) [@RN5536]; reverse - 806RB (5’ GGACTACNVGGGTWTCTAAT 3’) [@RN5537]. These primers also included the Illumina sequencing primer sequences attached to their 5’ ends. In the first amplification step, PCRs were carried out in a final volume of 12.5 μL, containing 1.25 μL of template DNA (1:2 diluted in PCR round 3; not diluted in PCR round 11), 0.5 μM of the primers, 3.13 μL of Supreme NZYTaq 2x Green Master Mix (NZYTech), and ultrapure water up to 12.5 μL. The fungal reaction mixture was incubated as follows: an initial denaturation at 95 ºC for 5 min, followed by 35 cycles of 95 ºC for 30 s, 49 ºC for 45 s, 72 ºC for 45 s, and a final extension step at 72 ºC for 7 min. The bacterial reaction mixture was incubated as follows: an initial denaturation at 95 ºC for 5 min, followed by 25 cycles of 95 ºC for 30 s, 47 ºC for 45 s, 72ºC for 45 s, and a final extension step at 72 ºC for 7 min. The oligonucleotide indices that are required for multiplexing different libraries in the same sequencing pool were attached in a second amplification step with identical conditions but only 5 cycles and 60 ºC as the annealing temperature. A negative control that contained no DNA (BPCR) was included in every PCR round to check for contamination during library preparation. Library size was verified by running the libraries on 2 % agarose gels stained with GreenSafe (NZYTech) and imaging them under UV light. Then, libraries were purified using the Mag-Bind RXNPure Plus magnetic beads (Omega Bio-tek), following the instructions provided by the manufacturer. Finished libraries were pooled in equimolar amounts according to the results of a Qubit dsDNA HS Assay (Thermo Fisher Scientific) quantification. The pool was sequenced in a fraction of a NovaSeq PE250 flow cell (Illumina) aiming for a total output of 2 gigabases. Any potential traces of adapter dimers were removed using Cutadapt v3.5 [@RN5538]. Quality of the sequencing data was assesed with the software FastQC [@RN5539].

*Processing of sequencing data and inference of ASVs*. Cutadapt v3.5 was used to trim off non-biological DNA sequences (primers, indices, and sequencing adapters). Then, amplicon reads were processed using the DADA2 pipeline [@RN5541] implemented in QIIME 2 (release 2023.7) [@RN5540]. The first step in the DADA2 pipeline was to remove PCR primers and filter reads according to their quality. After checking the read quality profiles, forward reads were truncated at position 175, and reverse reads at position 131. The next step was to denoise and infer Amplicon Sequence Variants (ASVs). Error rates were learned from the dataset to denoise, using the parametric error model implemented in DADA2. Dereplication of the dataset (i.e. combination of all identical reads into unique reads) was carried out to reduce computational effort. These dereplicated forward and reverse reads were used to infer ASVs with the core sample inference algorithm [@RN5541]. Subsequently, corresponding R1 and R2 reads were merged into pairs with a minimum overlap of 12 identical base pairs. The final step in the DADA2 pipeline was to reduce the impact of artefacts (chimaeras) in the dataset. The final output of the DADA2 pipeline was a table containing the count of reads of every observed ASV in each sample, as well as a table listing all ASVs and their corresponding representative sequences.

*Taxonomic assignment*. Taxonomic assignment of each ASV was performed using a pre-trained classifier of the UNITE reference database for fungi [@RN5550] (release May 2021); and of the SILVA reference database for bacteria [@RN5542] (release 138.1 August 2020). To compare the representative sequences of the ASVs to the reference database, and compute the taxonomic assignment,the algorithm sklearn [@RN5543] implemented in QIIME 2 was applied, setting a minimum confidence level of 0.8. Several filters were applied to exclude: (1) singletons (i.e. ASVs containing only one member sequence in the whole data set); (2) ASVs occurring at a frequency below 0.01 % in each sample, to reduce the mistagging bias [@RN5544]; (3) from the bacterial dataset, non-prokaryotic ASVs such as eukaryotic sequences of plastid [@RN5545] and mitochondrial [@RN5546] origin; (4) those assigned only at domain (Bacteria) or kingdom (Fungi) level; (5) ASVs present in the negative controls that were more abundant in the negative controls than in the samples, indicating contamination [@RN5547].

*Alpha rarefaction plots*. Alpha rarefaction plots showed that the sequencing depth to reach saturation was 30,000 reads for the fungi and 22,000 reads for the bacteria, above the sequencing depth of all plots for fungi (mean = 114,919 reads; min = 79,045; max = 157,110) and bacteria (mean = 113,341 reads; min = 84,094; max = 171,404).

> Rarefaction

## Statistical analysis

We performed all analyses using R version 4.3.1 [@RN5387] and the package *tidyverse* [@RN5387] for data manipulation and visualization. To describe soil temperatures as recorded by the dataloggers, we calculated the mean annual temperature, the mean diurnal range and the temperature annual range following the same methdology as the one applied to calculate bioclimatic indices bio1, bio2 and bio7 in WorldClim [@RN5064]. To assess the variability in soil physicochemical parameters, we conducted a Principal Component Analysis (PCA) as implemented in the R package *FactoMineR* [@RN3166], and a correlation analysis to detect auto-correlated soil variables. We used PERMANOVA as implemented in *vegan* [@RN5386] to test differences between habitat types in their soil variables and their composition of plant species, fungi ASVs and bacteria ASVs. We used Non-metric Multi-Dimensional Scaling (NMDS) ordination as implemented in *vegan* [@RN5386] to assess variation in taxonomic composition of plants (abundance values of species), fungi (presence/absence of ASVs) and bacteria (presence/absence of ASVs). We used environmental fitting as implemented in *vegan* [@RN5386] to check the correlation between the community ordination axes and the soil physicochemical properties. We used the NMDS multidimensional compositional distance from each urban plot (i.e. park, roadside, residential or industrial) to the centroid of the two reference non-urban habitats (i.e. forests and meadows) to calculate for each plot, taxonomic group and reference habitat a value of community novelty. We fitted a Generalized Mixed Model using the package *nlme* [@RN2333] to test the effect on comminity novelty of taxonomic group, habitat of reference (forest or meadow), urban habitat (park, roadside, residential or industrial) and soil physicochemical variables. The models inclused sampling plot as a random factor.

# Results

## Soil properties across habitats

Soil temperature records (**Figure 2A**) indicated that forest soils were relatively buffered in their range of variation, with a mean annual temperature of 11.4 ºC, a mean diurnal range of 1.72 ºC and an annual range of 4.74 ºC. The meadow and urban plots had more thermal oscillation, with a mean diurnal range above 4 ºC and an annual range above 10 ºC. PCA ordination of the soil physicochemical properties (**Appendix S1**) indicated that the first axis of variation, explaining 23% of the variance, ordered the soils according to their concentration of heavy metals. The variables with the highest contribution to this axis (> 10%) were the soil contents in Pb, Cu, Zn, Cd and Cr, which accounted for 66% of the variability of the first axis. This first axis did not separate urban from non-urban habitats: for example, vacant industrial lots had some of the highest concentration in heavy metals, but also some of the lowest. Likewise, forest plots were distributed across the first axis. The second axis of variation explained 21% of the variance and ordered soils according to their compaction and water retention capacities, with the highest contributing variables (> 10%) being organic matter content, bulk density, N concentration and pH, which made up 59% of the axis' variability. This axis did separate soils according to their habitat: for example, forest sites had high organic matter and N content, while industrial vacant lots had higher bulk density and pH. The third axis explained 10% of the variability and ordered soils according to their content in Mg, with the highest contributing variables (>10%) being Mg content, Mg/K ratio and conductivity, together explaining 50% of the axis. In this case, like for the first axis, there was no clear separation between habitats, with industrial sites having some of the highest and the lowest values. Correlation analysis showed strong correlations between the groups of variables contributing to the first, second and third axes (**Appendix S1**). Therefore, to avoid autocorrelation issues, for further analyses we kept the variables Pb content, organic matter content and Mg content (**Figure 2B**), which together provided a straightforward biological interpretation of patterns in soil composition. Of these values, as previously indicated by the PCA, organic matter was the only one with a clear pattern of variation: a decrease from the non-urban to the urban habitats. A PERMANOVA of the soil values indicated that there was not a significant difference between habitats in their soil physicochemical properties (100,000 permutations, R^2^ = 0.19, F = 1.15, p = 0.31).

```{r fig2, out.width = "450px", echo = FALSE, fig.cap = "Boxplot of soil properties accross habitats. (A) Temperatures recorded every hour in one sampling plot per habitat from 1 Jan 2024 to 31 Dec 2024. (B) Soil physicochemical properties. Values are shown for the three most explanatory and uncorrelated variables, as per a previous PCA analysis: soil Pb content, soil organic matter content, and soil Mg content."}
knitr::include_graphics("../results/figures/Fig2.png")
```

> Poner unidades de las variables en figura

## Plant, fungal and bacterial communities across habitats

The survey of plant diversity yielded 216 species, with the average number per plot being 26 (min = 10, max = 38). Of these species, 48% were only found in urban-habitats (parks, roadsides, residential lots and industrial lots), 27% occurred only in non-urban habitats (forests and meadows) and 25% were shared by both urban and non-urban habitats (**Fig3A**). On average, the habitat with the highest number of species per plot were the industrial lots (29), followed by residential lots (27), roadsides (24), meadows (23), parks (20) and forests (18). The most frequent species were *Holcus lanatus* (22 plots), *Plantago lanceolata* (20), *Helminthotheca echioides* (16), *Dactylis glomerata* (15) and *Lotus corniculatus* (14). PERMANOVA indicated that the plant species composition was significantly different accross habitats (100,000 permutations, R^2^ = 0.42, F = 3.48, p = < 0.001). NMDS ordination of the plant species composition (**Fig. 3D**) indicated that the plots belonging to the same habitat were relatively grouped together, with the main exception being one of the roadside plots. The ordination indicated that the major gradient in vegetation composition along the first axis was a separation between the forest plots and the rest, with the meadows being the closest to the forests. Environmental fitting of the Pb, organic matter and Mg vectors onto the ordination indicated that none of the soils factors had a significant correlation (R^2^ < 0.14, p > 0.05) with the composition of the plant communities.

The DNA metabarcoding of soil fungal communities resulted in 4,689 ASVs. Of these ASVs, 46% were only found in urban-habitats (parks, roadsides, residential lots and industrial lots), 36% occurred only in non-urban habitats (forests and meadows) and 18% were shared by both urban and non-urban habitats (**Fig3B**). The average number of ASVs per plot was 340 (min = 174, max = 587), with the habitat-averages ranging from residential vacant lots (415) to forests (393), meadows (336), parks (323), industrial vacant lots (300) and roadsides (276). The majority of ASVs were assigned to Ascomycota (56%), Basidiomycota (24%), Rozellomycota (10%) and Glomeromycota (4%). Of the different habitats, forests had the highest proportion of Basidiomycota (25% vs. < 17% in the other habitats) and the lowest proportion of Ascomycota (57% vs. > 64%). The proportion of Rozellomycota was similar across habitats (from 6% in meadows to 11% in parks). The proportion of Glomeromycota varied between 5% in meadows to 0.5% in forests. A PERMANOVA performed on the presence/absence of ASVs indicated significant differences in composition between habitats (100,000 permutations, R^2^ = 0.33, F = 2.39, p = < 0.001). NMDS ordination of the fungal ASV composition (**Fig. 3E**) behavead similarly to the ordination of plant communities, with plots tending to group by habitat, and the major compositional gradient reflecting the separation between the forest plots and the rest. Environmental fitting of the Pb, organic matter and Mg vectors onto the ordination indicated only organic matter content had a significant correlation (R^2^ = 0.46,  p < 0.001) with the composition of the fungal communities. Low organic matter content correlated with the communities from industrial vacant lots and roadsides (**Fig. 3E**).

For the bacterial communities, DNA metabarcoding yielded 16,306 ASVs. Of these ASVs, 44% were only found in urban-habitats (parks, roadsides, residential lots and industrial lots), 31% occurred only in non-urban habitats (forests and meadows) and 25% were shared by both urban and non-urban habitats (**Fig3C**). The average number of ASVs per plot was 1769 (min = 1478, max = 2074), with the habitat-averages ranging from roadsides (1848) to industrial lots (1794), residential lots (1792), meadows (1775), parks (1706) and forests (1697). The majority of ASVs were assigned to Proteobacteria (22%), Planctomycetota (15%), Actinobacteriota (15%) and Acidobacteriota (10%). Compared to other habitats, forests had higher proportions of Proteobacteria and Proteobacteria; a lower proportion of Actinobacteriota; and a similar proportion of Planctomycetota. A PERMANOVA performed on the presence/absence of ASVs indicated significant differences in composition between habitats (100,000 permutations, R^2^ = 0.32, F = 2.51, p = < 0.001). NMDS ordination of the bacterial ASV composition (**Fig. 3F**) resulted in less grouping by habitat than for the plants and the fungi, but still reflected a main composotional gradient from forests to industrial vacant plots. Like in the case of the fungal communities, environmental fitting of the Pb, organic matter and Mg vectors onto the ordination indicated only organic matter content had a significant correlation (R^2^ = 0.39,  p = 0.001) with the composition of the fungal communities. Low organic matter content correlated with the communities from industrial vacant lots and roadsides (**Fig. 3F**).

```{r fig3, out.width = "450px", echo = FALSE, fig.cap = "Plant, fungal and bacterial diversity of the sampling plots. Top panels show Venn plots for the shared numbers of (A) plant species, (B) fungal ASVs and (c) bacterial ASVs between urban (parks, roadsides, residential lots and industrial lots) and non-urban (forests, meadows) habitats. Bottom panels show  biplots of community composition ordinartion produced by Non-metric Multi-Dimensional Scaling (NMDS). NMDs were conducted for (A) plant survey (by species abundance), (B) the soil fungal survey (by ASV presence/absence) and (C) the soil bacterial survey (by ASV presence/absence). Colors indicate the habitat. In the biplots, arrows and labels indicate soil variables with a significant correlation with the composition of the communities. Only soil organic matter content and fungal-bacterial communities had a significant correlation."}
knitr::include_graphics("../results/figures/Fig3.png")
```

## Community novelty across habitats and taxonomic groups

We used the NMDS multidimensional compositional distance (calculate from the ordinations in **Fig. 3D-F**) from each urban plot (i.e. park, roadside, residential or industrial) to the centroid of the two reference non-urban habitats (i.e. forests and meadows) to calculate for each plot, taxonomic group and reference habitat a value of community novelty (**Fig. 4**). The novelty was significantly different between taxonomic groups (F = 45.88, p < 0.001), being higher for plants (2.45 ± 0.332 S.E.) than for fungi (2.15 ± 0.396 S.E.) and bacteria (1.56 ± 0.303 S.E.). The novelty was also significantly different depending on the reference non-urban habitat that was used (F = 199.08, p < 0.001), being higher compared to forests (2.60 ± 0.287 S.E.) than to meadows (1.51 ± 0.300 S.E.). Nonetheless, for each taxonomic group, there was a high positive correlation between the novelty compared against forests or meadows (r = 0.5 for plants; 0.7 for fungi; 0.9 for bacteria). The different urban habitats also differed significantly in their novelty (F = 9.51, p = 0.001), being higher for industrial vacant lots (2.46 ± 0.311 S.E.) than for roadsides (2.18 ± 0.367 S.E.), parks (1.95 ± 0.440 S.E.) and residential vacant lots (1.63 ± 0.307 S.E.). The soil physicochemical properties (Pb, organic matter, Mg) had no significant effect on community novelty (p > 0.05).

```{r fig4, out.width = "450px", echo = FALSE, fig.cap = "Community novelty for plants, fungi and bacteria. Community novelty was calculated as the multidimensional distance from each urban plot to the centroid of two reference non-urban habitats (forests, panels A-C-E; and meadows, panels B-D-F), as calculated by NMDS ordination. Novelty was calculated separately for plant (panels A-B), fungal (panels C-D) and bacterial communities (panels E-F)."}
knitr::include_graphics("../results/figures/Fig4.png")
```

## Soil drivers of novelty

# Discussion

The recorded flora represents just a subset of possible spontaneous vegetation because this study was carried on during limited time which covers just part of the growing season of the whole vegetation.

There are some other habitats in the city which were not included in the present study, e.g. wetlands, beach, cliffs.

# Data availability

The original datasets, as well as R code for analysis and creation of the manuscript can be accessed at the GitHub repository https://github.com/efernandezpascual/disturbance. Upon publication, a version of record of the repository will be deposited in Zenodo.

# References

::: {#refs}
:::

# Supplementary

**Appendix S1** Principal Component Analysis and correlation analysis of the soil physicochemical properties.