### 2025-03-17 README.Rmd
submission UFUG

---
title: "README.md"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

![](xove.JPG)

# Compositional novelty of plant, fungal and bacterial communities across urban habitats

This repository stores all the information related to the manuscript *Compositional novelty of plant, fungal and bacterial communities across urban habitats*, including the datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.

## Contents

This repository is organised following the advice of [Wilson et al. 2017](https://doi.org/10.1371/journal.pcbi.1005510) for recording and storing research projects.

The following materials are available in the folders of this repository:

* `data` Data files including the soil, plant, fungi and bacteria data.
* `src` Scripts in `R` language used to perform the analyses and create the figures.
* `results` Output of the `R` scripts, including figures and supplementary material.
* `doc` Files to create the [manuscript](https://github.com/efernandezpascual/xixon/blob/master/doc/manuscript.md) using `Rmarkdown`.

## Citation

Please cite the repository, datasets and article as:

> Fernández-Pascual E, Ferencova Z, González-García, Jiménez-Alfaro B. Compositional novelty of plant, fungal and bacterial communities across urban habitats