### 2025-07-07 doc/R0/title-page.Rmd
LANDUP R1

---
title: "Compositional novelty of plant, fungal and bacterial communities across urban habitats"
output: 
  bookdown::word_document2:
    reference_docx: template.docx
    number_sections: false
bibliography: refs.bib
csl: urban-forestry-and-urban-greening.csl
link-citations: yes
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",")
})
```

Running title: Compositional novelty across urban habitats

Eduardo Fernández-Pascual^1,a^ (<https://orcid.org/0000-0002-4743-9577>), Zuzana Ferencova^2^ (<https://orcid.org/0000-0002-0915-0916>), Víctor González-García^1^ (<https://orcid.org/0000-0002-8949-7943>), Borja Jiménez-Alfaro^1^ (<https://orcid.org/0000-0001-6601-9597>)

^1^ Biodiversity Research Institute (IMIB), University of Oviedo - CSIC - Principality of Asturias, Mieres, Spain; ^2^ Jardín Botánico Atlántico, Gijón/Xixón, Spain

Correspondence: Eduardo Fernández-Pascual, Instituto Mixto de Investigación en Biodiversidad, Campus de Mieres, Edificio de Investigación, 5ª planta, c/ Gonzalo Gutiérrez Quirós s/n, E-33600 Mieres, Spain. Email: [fernandezpeduardo\@uniovi.es](mailto:fernandezpeduardo@uniovi.es){.email}. Telephone: +34985104781.

# Author contributions

Eduardo Fernández-Pascual: Conceptualization; Methodology; Investigation; Data Curation; Formal Analysis; Visualization; Writing – Original Draft Preparation; Writing – Review & Editing. Zuzana Ferencova: Conceptualization; Methodology; Investigation; Data Curation; Writing – Review & Editing. Víctor González-García: Investigation; Visualization; Writing – Review & Editing; Borja Jiménez-Alfaro: Methodology; Writing – Review & Editing.

# Acknowledgements

María García Fernández, Sara Gutiérrez Cruz and Abelardo Sigüenza Solís helped with the field surveys. Thanks to AllGenetics & Biology SL (A Coruña, Spain) for DNA metabarcoding analysis and to INEA Laboratory (Valladolid, Spain) for soil physicochemical analysis. We are thankful to Fundación Biodiversidad for their support.

# Declaration of interest statement

We have nothing to declare.

# Funding

European Union - NextGenerationEU within the framework of the Recovery, Transformation and Resilience Plan (PRTR) of Ministry for the Ecological Transition and the Demographic Challenge (MITECO) of the Government of Spain —Project: Gijón Ecoresiliente (PRTR 0204103/010000). Ayuntamiento de Gijón/Xixón, Jardín Botánico Atlántico [SV-23-GIJON-JBA]. Grant "Laboratorio de Vegetación y Biodiversidad" (IDE/2024/000720, Principality of Asturias-Sekuens-EU-FEDER). Víctor González-García was supported by the Government of Asturias (Programa “Severo Ochoa” PA-22-BP21-098).

# Data availability statement

Original datasets, R code for analysis and Rmarkdown code for creation of the manuscript can be accessed at Zenodo: [DOI WILL BE INSERTED HERE].