### 2025-08-22 doc/R1/novelty-PCA.Rmd

---
title: "Untitled"
output: html_document
date: "2025-06-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Compositional novelty across habitats and taxonomic groups

We used the NMDS multidimensional compositional distance (calculated from the ordinations in **Fig. 3D-F**) from each of the urban plots (i.e., park, roadside, residential or industrial) to the centroid of the two reference pre-urban habitats (i.e., forests and meadows) to calculate a compositional novelty index (**Fig. 4**). Novelty was significantly different between taxonomic groups (F = 65.83, p \< 0.001), being higher for plants (1.55 ± 0.14 S.E.) than for fungi (1.20 ± 0.10 S.E.) and bacteria (1.01 ± 0.14 S.E.). Novelty was also significantly different depending on the pre-urban habitat that was used as a reference (F = 59.89, p \< 0.001), being higher compared to forests (1.40 ± 0.15 S.E.) than to meadows (1.10 ± 0.15 S.E.). Different urban habitats also differed significantly in their novelty (F = 8.03, p = 0.0028), being higher in industrial vacant lots (1.41 ± 0.14 S.E.) than in parks (1.26 ± 0.16 S.E.), roadsides (1.24 ± 0.17 S.E.) and residential vacant lots (1.09 ± 0.16 S.E.). The soil physicochemical properties (Pb, organic matter, Mg) had no significant effect on compositional novelty (p \> 0.05).

```{r fig4, echo = FALSE, fig.cap = "Compositional novelty for plants, fungi and bacteria. Compositional novelty was calculated as the NMDS multidimensional distance from each urban plot to the centroid of two reference pre-urban habitats (forests, panels A-C-E; and meadows, panels B-D-F). Novelty was calculated separately for plants (panels A-B), fungi (panels C-D) and bacteria (panels E-F)."}
knitr::include_graphics("../results/figures/Fig4-PCA.png")
```