### 2025-08-22 doc/R1/title-page.Rmd

---
title: "Compositional novelty of plant, fungal and bacterial communities across urban habitats"
output: 
  bookdown::word_document2:
    reference_docx: template.docx
    number_sections: false
bibliography: refs.bib
csl: landscape-and-urban-planning.csl
link-citations: yes
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",")
})
```

Running title: Compositional novelty across urban habitats

Eduardo Fernández-Pascual^1,2,a^ (<https://orcid.org/0000-0002-4743-9577>), Zuzana Ferencova^3^ (<https://orcid.org/0000-0002-0915-0916>), Víctor González-García^1,2^ (<https://orcid.org/0000-0002-8949-7943>), Borja Jiménez-Alfaro^1,2^ (<https://orcid.org/0000-0001-6601-9597>)

^1^ Biodiversity Research Institute (IMIB), University of Oviedo - CSIC - Principality of Asturias, Mieres, Asturias, Spain; ^2^ Department of Organismal and Systems Biology, University of Oviedo, Oviedo/Uviéu, Asturias, Spain; ^3^ Jardín Botánico Atlántico, Gijón/Xixón, Spain

Correspondence: Eduardo Fernández-Pascual, Instituto Mixto de Investigación en Biodiversidad, Campus de Mieres, Edificio de Investigación, 5ª planta, c/ Gonzalo Gutiérrez Quirós s/n, E-33600 Mieres, Spain. Email: [fernandezpeduardo\@uniovi.es](mailto:fernandezpeduardo@uniovi.es){.email}. Telephone: +34985104781.

# Author contributions

Eduardo Fernández-Pascual: Conceptualization; Methodology; Investigation; Data Curation; Formal Analysis; Visualization; Writing – Original Draft Preparation; Writing – Review & Editing. Zuzana Ferencova: Conceptualization; Methodology; Investigation; Data Curation; Writing – Review & Editing. Víctor González-García: Investigation; Visualization; Writing – Review & Editing; Borja Jiménez-Alfaro: Methodology; Writing – Review & Editing.

# Acknowledgements

María García Fernández, Sara Gutiérrez Cruz and Abelardo Sigüenza Solís helped with the field surveys. Thanks to AllGenetics & Biology SL (A Coruña, Spain) for DNA metabarcoding analysis and to INEA Laboratory (Valladolid, Spain) for soil physicochemical analysis.

# Declaration of interest statement

We have nothing to declare.

# Funding

The project is supported by the Biodiversity Foundation of the Ministry for the Ecological Transition and the Demographic Challenge (MITECO) of the Government of Spain, within the framework of the Recovery, Transformation and Resilience Plan (PRTR), funded by the European Union - NextGenerationEU. Ayuntamiento de Gijón/Xixón, Jardín Botánico Atlántico (SV-23-GIJON-JBA). Grant "Laboratorio de Vegetación y Biodiversidad" (IDE/2024/000720, Principality of Asturias-Sekuens-EU-FEDER). Víctor González-García was supported by the Government of Asturias (Programa “Severo Ochoa” PA-22-BP21-098).

# Data availability statement

Original datasets, R code for analysis and Rmarkdown code for creation of the manuscript can be accessed at Zenodo: [DOI WILL BE INSERTED HERE].