### 2025-08-22 results/supplementary/Supplementary-material-2.Rmd

---
title: "Supplementary material 2 Principal Component Analysis and correlation analysis of the soil physicochemical properties"
output:
  bookdown::pdf_document2:
    reference_docx: template.docx
    number_sections: false
    toc: false
link-citations: yes
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",")
})
```

```{r figS2a, out.width = "450px", echo = FALSE, fig.cap = "(A) Biplot of the first and second axes of a Principal Component Analysis of the soil physicochemical properties. Circles represent the sampling plots, coloured by habitat. Labeled arrows indicate the contribution of the soil variables to the axes."}
knitr::include_graphics("figS2a.png")
```

```{r figS2b, out.width = "450px", echo = FALSE, fig.cap = "(B) Biplot of the first and third axes of a Principal Component Analysis of the soil physicochemical properties. Circles represent the sampling plots, coloured by habitat. Labeled arrows indicate the contribution of the soil variables to the axes."}
knitr::include_graphics("figS2b.png")
```

```{r figS2c, out.width = "450px", echo = FALSE, fig.cap = "(C) Correlation heatmap of the soil variables."}
knitr::include_graphics("figS2c.png")
```