### 2025-08-22 results/supplementary/Supplementary-material-3.Rmd

---
title: "Supplementary material 3 Full results of the Linear Mixed Model"
output:
  bookdown::pdf_document2:
    reference_docx: template.docx
    number_sections: false
    toc: false
link-citations: yes
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",")
})
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
library(tidyverse)
load(file = "../model/m1.Rdata")
```

## Model summary

```{r message = FALSE, echo = FALSE, warning = FALSE}
library(nlme)
summary(m1)
```

## ANOVA table

```{r message = FALSE, echo = FALSE, warning = FALSE}
library(nlme)
anova.lme(m1)
```

## Diagnostic plots

```{r message = FALSE, echo = FALSE, warning = FALSE}
library(performance)
library(see)
check_model(m1)
```